---
title: "RMarkdown Lesson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Markdown text has formatting directives

# One hash is title header

## Two hashes is second level header

*This asterisk makes italicized text*

**Double asterisk makes bold text**

* create a list by making dashes
* apple
* banana
* kiwi
* these will be bulleted
+ this is a sub item
+ this is another sub item

1. Here is an ordered list
2. And here is another item
  + and a sub item
  + and another sub item

1. If you make a numbered list as all ones
1. R is smart enough
1. To put in the right numbers
1. This is useful for long lists
1. And reordering lists

_italicized...go to tools-> markdown quick reference to see all of the formatting options
an enter does not necessarily mean a new line_

**Add two empty lines if you want to make a new line**

### This is header 3


[to include links put link text here] 

in between dollar sign you can put LaTEX syntax

$\sum {n+1}$

Start a new chunk with the back ticks and r, and end it with three back ticks
command+opt+i adds a chunk for you!

```{r}
# This is a comment
x <- 4+4
x
```

Don't put in back to back r chunks

Putting the r in brackets tells it that you are writing r code and not python for example

```{r}
y <- 2^3
y

print(x+y)
```

