---
title: "Session 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(readr)
```



## Data Sources

### Read in the data

One nice feature of the read_csv (compared to the read.csv command) is that it can read in from a web url. Using a url instead of a file path make this much more reproducible.

```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

# Control shift M is shortcut for the pipe
```


## Clean and Reshape the data
### Remove unnecessary columns

```{r}
 catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)


# Or a shorter way
 catch_data <- catch_original %>% 
  select(-All, -notesRegCode)
 
 summary(catch_data)
```



### Check column typing
```{r}
summary(catch_data)

catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))


is.na(catch_data)
which(is.na(catch_clean$Chinook))  #tells you which observation/s are na 

# There is one obs that is an I instead of a one

catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook=="I", 1, Chinook)) %>%
  mutate(Chinook = as.numeric(Chinook))

# If the first statement is true, uses value of second argument, if false, uses value in the Chinook column


```
Oh no! Chinook is a character variable

### Reshape data

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(-c(Region, Year), names_to="species", values_to="catch")

head(catch_long)

catch_wide <- catch_long %>% 
  pivot_wider(names_from = species, values_from = catch)

head(catch_wide)
```
## Update value to thousands

```{r}
catch_long <- catch_long %>% 
  rename(catch_thousands = catch)
```

```{r}

catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)
head(catch_long)
```
```{r}
# Split/apply/combine all at once
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(catch_mean = mean(catch)) %>% 
  arrange(- catch_mean)

head(mean_region)


```
```{r}
n_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(n=n())

head(n_region)
```

```{r}

SSE_catch <- catch_long %>% 
  filter(Region=="SSE")

head(SSE_catch)
```

## Join to Regions dataset

```{r}
region_defs <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
select(code, mgmtArea)

catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

Two other very useful functions: separate and unite

```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAI-039"))

sites_df %>% 
  separate(site, c("island", "site_number"), "-")

```

```{r}
dates_df <- data.frame(year = c("1930",
                                "1930",
                                "1930"),
                       month = c("12",
                                "12",
                                "12"),
                       day = c("14",
                               "15",
                               "16"),
                       stringsAsFactors = FALSE)

dates_df %>% 
  unite(date, year, month, day, sep = "-") #put name of new column first
```
```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                  stringsAsFactors = FALSE)
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), names_to = "species", values_to = "catch") %>%
  mutate(catch = catch*1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  left_join(region_defs, by = c("Region" = "code"))
```
